### 2025-04-10 Paper/figures/Register_Based_Fragment.tex
Put diagrams into paper. Still need scaling fixes.

% Register based fragment
\begin{tikzpicture}

	\newcommand\drawThreadPhase[1]{
		\edef\phaseNum{#1}
% Thread\register layout
		\foreach \phaseRow in {0, ..., 7} {
			\foreach \phaseCol in {0, ..., 3}{
				\coordinate (C) at (\phaseCol * \registerWidth, -\phaseRow * \registerHeight);
				\pgfmathsetmacro\threadIndex{int(\phaseRow * 4 + \phaseCol)}
				\def\chunkColor{\getPhaseColor{\phaseNum}}
				\filldraw[fill=\chunkColor, draw=black] (C) rectangle +(\registerWidth, -\registerHeight) node[midway, font=\small] {\texttt{T\threadIndex}};
			}
		}
	}

	\drawFragment{ \drawThreadPhase }

\end{tikzpicture}