### 2025-04-10 Paper/figures/Single_TC_MMA.tex

% Basic TC MMA
\begin{tikzpicture}
	\newcommand\fm{3}
	\newcommand\fn{2}
	\newcommand\fk{3}

		% Draw B
	\pgfmathsetmacro\bdXShift{1.0}
	\draw[style=fragment,draw=BFragmentColor, ultra thick] (0, 0) rectangle +(\fn, -\fk);
	\matrix (m) [font=\tiny, xshift=\bdXShift cm, yshift=-1.5 cm, 
	matrix of math nodes,
	row sep=0.0, column sep=-0.5em,
	nodes={anchor=center}] {
		C_{1,1} & \cdots & C_{8,1} \\
		C_{1,2} & \cdots & C_{8,2} \\
		\vdots & \ddots & \vdots \\
		C_{1,15} & \cdots & C_{8,15} \\
		C_{1,16} & \cdots & C_{8,16} \\
	};

		% Draw D
	\pgfmathsetmacro\dXShift{0}
	\pgfmathsetmacro\dYShift{-\whiteSpace - \fk}
	\draw[style=fragment,draw=DFragmentColor, xshift=\dXShift cm, yshift=\dYShift cm, ultra thick] (0, 0) rectangle +(\fn, -\fm);
	\pgfmathsetmacro\adYShift{-4.75}
	\matrix (m) [font=\tiny, xshift=\bdXShift cm, yshift=\adYShift cm, 
	matrix of math nodes,
	row sep=0.0, column sep=-0.5em,
		%left delimiter={[}, right delimiter={]},
	nodes={anchor=center}] {
		D_{1,1} & \cdots & D_{1,8} \\
		D_{2,1} & \cdots & D_{2,8} \\
		\vdots & \ddots & \vdots \\
		D_{15,1} & \cdots & D_{15,8} \\
		D_{16,1} & \cdots & D_{16,8} \\
	};

		% Draw A
	\pgfmathsetmacro\aXShift{-\whiteSpace - \fk}
	\pgfmathsetmacro\aYShift{-\whiteSpace - \fm}
	\draw[style=fragment,draw=AFragmentColor, xshift=\aXShift cm, yshift=\aYShift cm, ultra thick] (0, 0) rectangle +(\fk, -\fm);
	\matrix (m) [font=\tiny, xshift=-1.75 cm, yshift=\adYShift cm, 
	matrix of math nodes,
	row sep=0.0, column sep=-0.5em,
	nodes={anchor=center}] {
		Q_{1,1} & Q_{1,2} & \cdots & Q_{1,16} \\
		Q_{2,1} & Q_{2,2} & \cdots & Q_{2,16} \\
		\vdots & \vdots & \ddots & \vdots \\
		Q_{15,1} & Q_{15,2} & \cdots & Q_{15,16} \\
		Q_{16,1} & Q_{16,2} & \cdots & Q_{16,16} \\
	};

\end{tikzpicture}