### 2025-05-01 Paper/figures/Single_TC_MMA.tex

% Basic TC MMA
\begin{tikzpicture}
	\newcommand\fm{2.5}
	\newcommand\fn{1.25}
	\newcommand\fk{\fm}
        \pgfmathsetmacro\whiteSpace{0.3}

    \def\matrixLabelShift{0}
    \tikzstyle matrixLabelStyle=[midway, text=white, font=\small, yshift=-\matrixLabelShift mm, xshift=\matrixLabelShift mm]
    \tikzstyle matrixTextStyle=[text=white, font=\tiny]
    \tikzstyle lineStyle=[draw=white]

		% Draw B
	\pgfmathsetmacro\bdXShift{\fn / 2}
	\pgfmathsetmacro\bdYShift{\fm / 2}
	\fill[style=fragment,fill=BFragmentColor] (0, 0) rectangle +(\fn, -\fk) node[style=matrixLabelStyle] {\qrf};
	\matrix (m) [style=matrixTextStyle, xshift=\bdXShift cm, yshift=-\bdYShift cm, 
	matrix of math nodes,
	row sep=1.5 cm, column sep=0.0 cm,
	nodes={anchor=center}] {
		p_{1,1} & p_{8,1} \\
		p_{1,16} & p_{8,16} \\
	};
    \draw[style=lineStyle] (m-1-1.south) -- (m-2-1.north);
    \draw[style=lineStyle] (m-1-2.south) -- (m-2-2.north);
    %\node[style=matrixLabelStyle] at (0 cm, 0 cm) {\qrf};

		% Draw D
	\pgfmathsetmacro\dXShift{0}
	\pgfmathsetmacro\dYShift{-\whiteSpace - \fk}
	\fill[style=fragment,fill=DFragmentColor, xshift=0 cm, yshift=\dYShift cm] (0, 0) rectangle +(\fn, -\fm) node[style=matrixLabelStyle] {\arf};
	\pgfmathsetmacro\adYShift{\dYShift - 0.5 * \fm}
	\matrix (m) [style=matrixTextStyle, xshift=\bdXShift cm, yshift=\adYShift cm, 
	matrix of math nodes,
	row sep=1.5 cm, column sep=0 cm,
	nodes={anchor=center}] {
		a_{1,1} & a_{1,8} \\
		a_{16,1} & a_{16,8} \\
	};
    % Draw all interconnecting lines
    %\draw (m-1-1.east) -- (m-1-2.west);
    %\draw (m-2-1.east) -- (m-2-2.west);
    %\draw (m-1-1.south) -- (m-2-1.north);
    %\draw (m-1-2.south) -- (m-2-2.north);
    \pgfmathsetmacro\matrixLabelYOffset{-\fm - \whiteSpace}
    ;

		% Draw A
	\pgfmathsetmacro\aXShift{-\whiteSpace - \fk}
	\pgfmathsetmacro\aYShift{-\whiteSpace - \fm}
	\pgfmathsetmacro\matXShift{-\fk / 2 - \whiteSpace}

	\fill[style=fragment,fill=AFragmentColor, xshift=\aXShift cm, yshift=\aYShift cm] (0, 0) rectangle +(\fk, -\fm) node[style=matrixLabelStyle] {\prf};
	\matrix (m) [style=matrixTextStyle, xshift=\matXShift cm, yshift=\adYShift cm, 
	matrix of math nodes,
	row sep=1.5 cm, column sep=1.0 cm,
	nodes={anchor=center}] {
		p_{1,1} & p_{1,16} \\
		p_{16,1} & p_{16,16} \\
	};
    \draw[style=lineStyle] (m-1-1.east) -- (m-1-2.west);
    \draw[style=lineStyle] (m-2-1.east) -- (m-2-2.west);
    \pgfmathsetmacro\matrixLabelXOffset{-\fk - \whiteSpace}
    %\node[style=matrixLabelStyle] at (\matrixLabelXOffset cm, \matrixLabelYOffset cm) {\prf};

%\begin{scope}[xshift=1.625 cm, yshift=0 cm, node distance=0.75 cm]
%    \node[anchor=north west] at (0, 0) {\small{\underline{Steps (GPU Core):}}};
%    \node[anchor=north west] at (0, -0.75) {\small{1. $s_i = \sum\limits_k{p_{ik}^2}$ (CUDA)}};
%    \node[anchor=north west] at (0, -1.5) {\small{2. $\arf = \prf \times \qrf + \arf$ (Tensor)}};
%    \node[anchor=north west] at (0, -2.25){\small{3. $dist_{ij}^2 = -2 \times a_{ij} + s_i + s_j$ (CUDA)}};
%    \end{scope}
\end{tikzpicture}